---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.path = "man/figure/"
)
```

# tsibble

[![Travis-CI Build Status](https://travis-ci.org/earowang/tsibble.svg?branch=master)](https://travis-ci.org/earowang/tsibble)
[![Coverage Status](https://img.shields.io/codecov/c/github/earowang/tsibble/master.svg)](https://codecov.io/github/earowang/tsibble?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/tsibble)](https://cran.r-project.org/package=tsibble)

## Overview

The **tsibble** package provides a data class of `tbl_ts` to manage temporal data frames in a tidy and modern way. A *tsibble* consists of a time index, keys and other measured variables in a data-centric format, which is built on top of the *tibble*.

## Installation

You could install the development version from Github using

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("earowang/tsibble", build_vignettes = TRUE)
```

## Get started

### Coerce to a tsibble with `as_tsibble()`

The `weather` data included in the package `nycflights13` is used as an example to illustrate. The "index" variable refers to the `time_hour` containing the date-times, and the "key" is the `origin` as weather stations. **The key together with the index uniquely identifies each observation**, which defines a valid *tsibble*. Others can be considered as measured variables.

```{r nycflights13, message = FALSE}
library(tsibble)
weather_ts <- as_tsibble(nycflights13::weather, origin, index = time_hour)
weather_ts
```

The **key** is not constrained to a single variable, but expressive for nested and crossed data structures. See `vignette()` for details.

### Summarise over calendar periods with `tsummarise()`

The common *dplyr* verbs, such as `mutate()`, `select()`, `filter()`, and `arrange()`, also work with the tsibble, except for `summarise()`. Because we have `tsummarise()` instead to aggregate interested variables over calendar periods.

```{r tsummerise}
weather_ts %>%
  group_by(origin) %>%
  tsummarise(
    year_month = yearmth(time_hour), # monthly aggregates
    avg_temp = mean(temp, na.rm = TRUE),
    ttl_precip = sum(precip, na.rm = TRUE)
  )
```

The `tsummerise` goes hand in hand with the index functions including `as.Date()`, `yearmth()`, `yearqtr()`, and `year()`.

## Related work

* [tibbletime](https://github.com/business-science/tibbletime)
* [tsbox](https://github.com/christophsax/tsbox)
