---
output:
  github_document:
    html_preview: false
    includes:
      before_body: header.md
    toc: true
    toc_depth: 3
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.path = "man/figure/"
)
set.seed(42)
options(tibble.print_min = 5)
```

## Installation

You could install the development version from Github using

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("earowang/tsibble", build_vignettes = TRUE)
```

## Get started

### Coerce to a tsibble with `as_tsibble()`

The `weather` data included in the package `nycflights13` is used as an example to illustrate. The "index" variable refers to the `time_hour` containing the date-times, and the "key" is the `origin` as weather stations. **The key together with the index uniquely identifies each observation**, which defines a valid *tsibble*. Others can be considered as measured variables.

```{r nycflights13, message = FALSE}
library(tsibble)
weather_ts <- as_tsibble(nycflights13::weather, origin, index = time_hour)
weather_ts
```

The **key** is not constrained to a single variable, but expressive of nested and crossed data structures. See `?tsibble` and `vignette("intro-tsibble")` for details.

### `tsummarise()` to summarise over calendar periods

A new verb `tsummarise()` is here to aggregate interested variables over calendar periods. The `tsummarise` goes hand in hand with the index functions including `as.Date()`, `yearmth()`, and `yearqtr()`, as well as other friends from *lubridate*, such as `year()` and `ceiling_date()`. For example, it would be of interest in computing average temperature and total precipitation per month, by applying the `yearmth()` to the hourly time index.

```{r tsummarise}
weather_ts %>%
  group_by(origin) %>%
  tsummarise(
    year_month = yearmth(time_hour), # monthly aggregates
    avg_temp = mean(temp, na.rm = TRUE),
    ttl_precip = sum(precip, na.rm = TRUE)
  )
```

The `tsummarise()` can be a useful function for regularising an irregular tsibble.

### `fill_na()` to turn implicit missing values into explicit missing values

Often there are implicit missing cases in temporal data. If the observations are made at regular time interval, we'd like to turn these implicit missings to be explicit. The `fill_na()` function not only extends the index and key to make the `NA`s present, but also provides a consistent interface to replace these `NA`s using a set of name-value pairs.

```{r fill-na}
nr <- nrow(weather_ts)
# randomly remove 20% of the observations
weather_na <- slice(weather_ts, sample(nr, size = nr * 0.8))
# replace NA with either functions or values for each group
weather_na %>%
  group_by(origin) %>%
  fill_na(
    year = lubridate::year(time_hour),
    temp = dplyr::lag(temp, na.rm = TRUE),
    precip = 0
  )
```

The missing values of the *year*, *temp* and *precip*, are supplied by the year extracted from the *time_hour*, the previous hour's temperature, and a single number of 0 respectively. The rest of untouched variables simply leave `NA` as is.

### Window functions applied to a tsibble: `slide()`, `tile()`, `stretch()`

Time series data commonly get involved in moving window calculations. A set of verbs provided in the *tsibble* allow for different variations of moving windows:

* `slide()`: sliding window with overlapping observations.
* `tile()`: tiling window without overlapping observations.
* `stretch()`: fixing an initial window and expanding more observations.

For example, a moving average of window size 3 is carried out on hourly temperatures for each group (*origin*).

```{r slide}
weather_ts %>% 
  select(origin, time_hour, temp) %>% 
  group_by(origin) %>% 
  mutate(temp_mv = slide(temp, mean, size = 3))
```

It can be noticed that the common *dplyr* verbs, such as `summarise()`, `mutate()`, `select()`, `filter()`, and `arrange()`, work with the tsibble. 

## Related work

* [tibbletime](https://github.com/business-science/tibbletime): time-aware tibbles. We have different APIs and thinking about temporal data.
* [padr](https://github.com/EdwinTh/padr): padding of missing records in time series. We do more.
