---
title: "Introduction to tsibble"
author: "Earo Wang"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to tsibble}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The **tsibble** package expands [tidyverse](https://www.tidyverse.org) to temporal-context data. Built on top of the [tibble](http://tibble.tidyverse.org), a tsibble (or `tbl_ts`) is a data-centric format, where time indices are preserved as the essential component and heterogeneous data structures are made possible. Beyond the tibble-like representation, a grammar is introduced to impose additional and informative structures on the tsibble, which is referred to as "key" variables. Multiple keys separated by a vertical bar (`|`) or a comma (`,`) are expressive of nested or crossed variables. This binds [hierarchical and grouped time series](http://pkg.earo.me/hts/) together into the tsibble framework.

```{r initial, echo = FALSE, cache = FALSE, results = 'hide'}
knitr::opts_chunk$set(
  warning = FALSE, message = FALSE, echo = TRUE,
  fig.width = 7, fig.height = 6, fig.align = 'centre'
)
options(tibble.print_min = 5)
```

The `as_tsibble()` is an S3 method to coerce other objects to a tsibble. An object that a vector/matrix underlies, such as `ts`, `mts`, or `hts`, can be automated to a tsibble using `as_tsibble()` without any specification. If it is a tibble or data frame, the `as_tsibble()` requires a little more setup in order to identify the index and key variables.

## Index and key(s)

```{r weather}
library(tsibble)
weather <- nycflights13::weather %>% 
  select(origin, time_hour, temp, humid, precip)
weather
```

```{r weather-ts}
weather_tsbl <- as_tsibble(weather, origin, index = time_hour)
weather_tsbl
```

```{r flights}
library(hms)
library(lubridate)
flights <- nycflights13::flights %>%
  mutate(sched_date_time = time_hour + minutes(minute)) %>%
  select(
    flight, origin, dest, sched_date_time, 
    dep_delay, arr_delay, air_time, distance
  )
```

```{r flights-ts}
flights_tsbl <- flights %>%
  as_tsibble(flight, origin, dest, index = sched_date_time, regular = FALSE)
flights_tsbl
```

```{r tourism}
tourism <- as_tsibble(tourism, Region | State, Purpose, index = Quarter)
```

## Interval
